---
title: "p8105_hw3_LC4052"
output: github_document
Author: "Linqi Chu"
date: "2025-10-06"
---

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
library(dplyr)

```


# Problem 1

## Import data

```{r}
library(p8105.datasets)
data("instacart")
```

## Description of variables

The dataset includes`r nrow(instacart)` observations and `r ncol(instacart)` variables, including `r names(instacart)`.

There are `r n_distinct(instacart$product_id)` distinct product ids and `r n_distinct(instacart$product_name)` product names. Departments contain `r unique(instacart$department)`.

## Number of aisles

```{r}
instacart |>
  count(aisle) |>
  arrange(desc(n)) |>
  slice(1) |>
  pull(aisle)

```
There are `r n_distinct(instacart$aisle` aisles. "fresh vegetables" are the most items ordered from.

## Plot of the number of items ordered in each aisle

```{r}
instacart |>
  count(aisle) |>
  filter(n > 10000) |>
  ggplot(aes(x = reorder(aisle, n), y = n)) +
  geom_col(fill = "steelblue", alpha = 0.8) +
  coord_flip() +
  labs(
    title = "Number of Items Ordered in Each Aisle",
    subtitle = "Aisles with more than 10,000 items ordered",
    x = "Aisle",
    y = "Number of Items Ordered"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma)

```

## Table of the three most popular items

```{r}
instacart |>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle, product_name) |>
  summarise(order_count = n()) |>
  group_by(aisle) |>
  slice_max(order_count, n = 3) |>
  arrange(aisle, desc(order_count))
 
```

## Table of the mean hour

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  mutate(
    day_of_week = factor(order_dow,
                       levels = 0:6,
                       labels = c("Sunday", "Monday", "Tuesday", "Wednesday",
                                  "Thursday", "Friday", "Saturday")),
    hour_of_day = order_hour_of_day
  ) |>
  group_by(product_name, day_of_week) |>
  summarize(mean_hour = round(mean(hour_of_day), 1)) |>
  pivot_wider(
    names_from = day_of_week,
    values_from = mean_hour
  )

```


# Problem 2

## Import and clean data

```{r}
# Import data
zipcodes_df = 
  read_csv("./zillow_data/Zip Codes.csv",
             na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  select(county,zip_code, everything())


# Clean the zip_codes file
zipcodes_df = 
  read_csv("./zillow_data/Zip Codes.csv",
             na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  select(county,zip_code, everything(), -file_date)

names(zipcodes_df)


# Clean the zip_zori file
zori_df = 
  read_csv("./zillow_data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv",
             na = c("NA", ".", "")) |>
  pivot_longer(
    cols = 10 : 125,
    names_to = "file_date", 
    values_to = "rent_price") |>
    mutate(file_date = as.character(file_date)) |>
  rename(county = CountyName, zip_code = RegionName) |>                       
  mutate(county = sub(" County$", "", county)) |> 
  arrange(zip_code) |>
  select(county,zip_code,file_date, rent_price, everything())


# Combine data
combined_data = 
  full_join(zipcodes_df, zori_df, by = c("county","zip_code")) |>
  select(county,zip_code,file_date, rent_price, neighborhood, everything())

names(combined_data)

```
## number of ZIP codes

```{r}
combined_data |>
  count(zip_code)|>
  filter(n == 116) |>
  nrow()

combined_data |>
  count(zip_code)|>
  filter(n < 10) |>
  nrow()

```
146 ZIP codes are observed 116 times.
171 ZIP codes are observed fewer than 10 times.


## Frequency differences of ZIP codes

```{r}
zip_freq =
  combined_data |>
  mutate(file_date = as.Date(file_date)) |>
  group_by(zip_code) |>
  summarize(
    n_months = n_distinct(format(file_date, "%Y-%m")),
    zip_count = n()
  )

ggplot(zip_freq, aes(x = n_months)) +
  geom_histogram(binwidth = 3, fill = "steelblue", color = "white") +
  labs(
    title = "Distribution of ZIP Code Observation Frequency",
    x = "Number of Months Observed",
    y = "Count of ZIP Codes"
  )
```
There are some reasons why some ZIP codes are observed rarely and others observed in each month:




## Table showing the average rental price in each borough and year (not month). Comment on trends in this table.

```{r}





```

```{r}
ggplot(zip_freq, aes(x = avg_rent)) +
  geom_histogram(binwidth = 500, fill = "steelblue", color = "white") +
  labs(
    title = "Distribution of ZIP Codes by Average Rent",
    x = "Average Rent ($)",
    y = "Number of ZIP Codes"
  ) +
  theme_minimal()
```


Make a plot showing NYC Rental Prices within ZIP codes for all available years. Your plot should facilitate comparisons across boroughs. Comment on any significant elements of this plot.

```{r}







```


 
Compute the average rental price within each ZIP code over each month in 2023. Make a reader-friendly plot showing the distribution of ZIP-code-level rental prices across boroughs; put differently, your plot should facilitate the comparison of the distribution of average rental prices across boroughs. Comment on this plot.

```{r}

```






Combine the two previous plots into a single graphic, and export this to a results folder in your repository

```{r}

```




# Problem 3

```{r}

```


```{r}

```










